A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-11-24 01_Introduction.Rmd
decided to start with a really introductory paragraph that gets explained/fleshed oout in the following sections

The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017]. Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live. Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002]. *Cultural Transmission Theory* is concerned with this exchange from the point of view of Cultural Evolution.

>>>>

This articles employs some of the methods that arose within archaeology's contemplation of *Cultural Transmission Theory* to trace the development of four major bronze age burial traditions in Europe. The cultural-historical evolution of archaeological cultures and ethnic groups during the european bronze age is complex and it's reconstruction an ongoing endeavour of archaeology. Within this history of constant emergence, displacement and demise many different forms of burial traditions were practiced. Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*. Bronze age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains. Openly available bulk radiocarbon data now allows to reconstruct the trends within the distribution of these complex behaviours to test classical archaeological narratives about their respective prevalence. The comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.

The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017]. Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live. Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
